Environmental and geographic variables are effective surrogates for genetic variation in conservation planning

### 2017-03-14 code/rmarkdown/article.Rmd
shorten article

---
title: "Environmental and geographic variables are effective surrogates for genetic variation in conservation planning"
author:
  - name: Jeffrey O. Hanson
    affiliation: a,1
  - name: Jonathan R. Rhodes
    affiliation: b
  - name: Cynthia Riginos
    affiliation: a
  - name: Richard A. Fuller
    affiliation: a
address:
  - code: a
    address: School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
  - code: b
    address: School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
corresponding_author:
  - code: 1
  - text: "To whom correspondence should be addressed. E-mail: jeffrey.hanson@uqconnect.edu.au"
author_contributions: |
  All authors developed the study and drafted the manuscript. JOH performed all analyses.
lead_author_surname: "Hanson"
abstract: |
  Protected areas buffer species from anthropogenic threats and provide places for the processes that generate and maintain biodiversity to continue. However, genetic variation, the raw material for many evolutionary processes, is difficult to capture in conservation planning, not least because they require considerable resources to obtain and analyze. Here we show that freely available environmental and geographic distance variables are highly effective surrogates in conservation planning for representing adaptive and neutral intra-specific genetic variation. We obtained occurrence and genetic data from the IntraBioDiv project for 27 alpine plant species collected over the European Alps using a gridded sampling scheme. For each species, we identified loci that were potentially under selection using outlier loci methods, and mapped their main gradients of adaptive and neutral genetic variation across the grid cells. We then used the cells as planning units to generate prioritize protected area acquisitions. First, we verified that the spatial patterns of environmental and geographic variation were correlated, respectively, with adaptive and neutral genetic variation. Second, we showed that these surrogates can predict the proportion of genetic variation secured in randomly generated solutions. Finally, we discovered that solutions based only on surrogate information secured substantial amounts of adaptive and neutral genetic variation. Our work paves the way for widespread integration of surrogates for genetic variation into conservation planning.
significance: |
  To protect biodiversity for the long term, nature reserves and other protected areas need to represent a broad range of different gentic types. However, genetic data are expensive and time consuming to obtain. Here we show that freely available environmental and geographic variables can be used as effective surrogates for genetic data in conservation planning. This means that conservation plannners can, with some confidence, design protected area systems to represent genetic diversity without investing in expensive programmes to obtain and analyze genetic data.
acknowledgements: |
  JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF has an Australian Research Council Future Fellowship. This work was supported by the Centre of Biodiversity and Conservation Science (CBCS).
keywords:
  - conservation
  - biodiversity
  - AFLP
  - genetic diversity
pnas_type: pnasresearcharticle
bibliography: references.bib
csl: reference-style.csl
output:
  rticles::pnas_article:
    includes:
      in_header: [preamble.tex]
---

<!-- NOTES -->
<!-- 150 word limit for abstract -->
<!-- 5000 word length limit -->

@@

```{r,include=FALSE}
# load .rda file
checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
try(session::restore.session('data/final/results.rda'))

@@

Broadly speaking, though the strength of natural selection is a continuous force, genetic variation is commonly classified as either adaptive or neutral (reviewed in [@r480]). Adaptive genetic variation is associated with loci that significantly affect fitness. Typically, such "adaptive" loci are detected due to anomalous patterns among individuals (eg. [@r456; @r487]), though phenotypic-genetic associations are more likely to identify functionally important genomic regions for individual species (eg. [@r500]). By representing the full range of adaptive variation, protected areas can help ensure that populations with particularly beneficial adaptations are not lost ([@r3; @r43]). In contrast, neutral genetic variation is associated with loci that do not significantly affect fitness, but instead reflect the evolutionary history of different populations. By representing the full range of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity ([@r67]). Thus optimally sited protected area networks would aim to represent both adaptive and neutral patterns of genetic variation ([@r67; @r3; @r43]).

Recently, the use of surrogate data has been proposed to generate prioritizations that sufficiently capture intra-specific genetic variation ([@r2; @r200])--without needing to directly utilize genetic data to achieve this. Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been proposed to guide reserve selection as surrogates for adaptive genetic variation (eg. [@r202; @r2]). This is on the basis that sampling a broad range of environmental conditions across species' distributions should also capture more adaptive variation. However, the effectiveness of environmental surrogates in representing adaptive genetic variation remains unverified. Neutral genetic variation arizes from a reduction in gene flow between individuals. Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood (reviewed in [@r501]). Potential surrogates for neutral variation have been based on variables that predict the levels of connectivity between planning units. For instance, isolation by distance (IBD; [@r259]) predicts that populations located further apart will experience less gene flow, and in turn, share fewer neutral loci. Based on this idea, geographic surrogates based on distance have been found to improve conservation prioritizations in an island network ([@r200]). However this remains untested in spatially contiguous systems where connectivity is complicated by many additional factors (eg. habitat; [@r497])--the typical situation in most conservation planning exercises.

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(7.1*0.9), out.width='3.42in', out.height='4in', fig.cap='The proportion of adaptive and neutral genetic variation secured for each species in solutions based on different targets and scenarios. Box plots show the 25th percentile, median, and 75th percentile. Whiskers show the data at 1.5 $\times$ the inter-quantile range. Points show data outside the whiskers. Red lines show the target proportion of genetic variation.'}
# prepare data for plotting
scenario.SDF <- scenario.DF %>%
  group_by(prioritization, Context, Metric, Species) %>%
  summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
  data.frame()

# make plot
suppressWarnings({
  ggplot(aes(x=prioritization,y=genetic.held),
    data=scenario.SDF) +
    geom_boxplot(position=position_dodge(0.9)) +
    scale_fill_manual(name='prioritization',
      values=c('grey80','grey50','grey20')) +
    geom_abline(slope=0, color='red', linetype='dashed',
      intercept=raptr.params.LST[[MODE]]$scenario.analysis$genetic.target) +
    ylab('Genetic variation secured (%)') +
    xlab('') +
    scale_y_continuous(breaks=c(0,0.25,0.5,0.75,1)) +
    theme_classic() +
    theme(strip.background = element_rect(fill='grey20'),
      strip.text = element_text(color='white', size=9),
      panel.border = element_rect(color='black', fill=NA),
      axis.text.x=element_text(angle=45, hjust=1, vjust=1, size=8),
      axis.text.y=element_text(size=8)) +
    facet_grid(Context ~ Metric)
})
```

The proportion of genetic variation secured in a prioritization depended on the targets used to generate the prioritization (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$p.value[[2]])`)$. Most notably, prioritizations based on the surrogate variables represented significantly more genetic variation than amount-based prioritizations (`r round(mean(filter(scenario.sub.DF,prioritization=='Amount & surrogate\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,prioritization=='Amount & surrogate\ntargets')$genetic.held*100),2)` SD overall genetic variation secured versus `r round(mean(filter(scenario.sub.DF,prioritization=='Amount targets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,prioritization=='Amount targets')$genetic.held*100),2)` SD; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$), and were not distinguishable from prioritizations based on measured genetic data (`r round(mean(filter(scenario.sub.DF,prioritization=='Amount & genetic\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,prioritization=='Amount & genetic\ntargets')$genetic.held*100),2)` SD; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$). These results indicate that environmental and geographic surrogates in this case far outperform traditional amount-based conservation planning, and perform as well as a conservation plan based directly on measured genetic data.

@@

<!-- surrogate/genetic held correlations -->
Our results demonstrate that environmental and geographic variables are effective surrogates for most species (Figure 2). Despite this, for a few species, geographic distance was a suprisingly poor surrogate for neutral genetic variation (eg. _Gypsophila repens_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`). Indeed, geographic distance is expected to be a poor surrogate for connectivity when species deviate from isolation by distance. For example, species can utilize disersal modes that maintain high connectivity between distant areas (eg. wind dispersing plants; [@r451]), and connectivity between adjacent populations can be disrupted by landscape features ([@r497]). Overall, such species were the exception in our analysis, and both surrogates performed moderately well for most species.

<!-- suitable targets for surrogates -->
Our results suggest that environmental and geographic surrogates can be used to capture genetic variation in prioritizations, the next question is: what percentage of surrogate variation should planners target to preserve genetic variation? Firstly, our results suggest that surrogate-based targets should be used with amount-based targets. For all species, it was possible to generate solutions that secured a high proportion of the surrogate variables (> 90 %) and only a small proportion of genetic variation (< 20 %; Figures S56--S57). By using both amount- and surrogate-based targets, planners may avoid such outcomes. Secondly, geographic surrogates seem to need higher targets than environmental surrogates to capture genetic variation. In our analysis, neutral genetic variation accumulated much more slowly with geographic representation than adaptive variation accumulated with environmental representation (Figures 2a--2b and S56--S57). To secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` % of the species' neutral genetic variation, prioritizations needed to sample `r round(mean(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` SD of the variation in geographic position among the planning units occupied by the species. On the other hand, to secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` % of each species' adaptive genetic variation, prioritizations only needed to sample `r round(mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` SD of the environmental variation in planning units occupied by the species. In many individual cases, higher targets may be needed to ensure that prioritizations secure a representative sample of the genetic variation for all of the species in a planning exercise.

<!-- adaptive genetic variation secured in prioritizations-->
Our results suggest that conservation planners can secure a reasonably representative sample of the intra-specific adaptive genetic variation using only amount-based targets. For most species, the spatial patterns of adaptive genetic variation clustered into a few distinct groups (eg. _Carex semipervens_ and _Cerastium uniflorum_; Figures S33 and S34 respectively). As a consequence, a prioritization would only need to select a single planning unit in each cluster to secure a representative sample. This mechanism explains why adaptive variation accumulated much more quickly than neutral genetic variation in the randomly generated prioritizations. Additionally, this mechanism also explains why the single-species prioritizations generated using only amount targets were often able to achieve high representation--they were essentially a random selection and so frequently contained planning units in different parts of the species' ranges (Figure 4a). Overall--and contrary to our original hypotheses--it seems that conventional reserve selection methods might often secure a representative sample of species' adaptive genetic variation.

<!--  limitations/ weaknesses -->
There are several limitations associated with our analysis. Firstly, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in regional conservation planning scenarios (1--10 km$^2$). We used this resolution because the genetic data were collected at this scale. Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis by artificially introducing additional spatial auto-correlation. Secondly, we used geographic distances as surrogates for neutral genetic variation, but connectivity in this landscape may be better described using distances that incorporate information on topography. However, such distances require species-specific scaling to adequately describe their connectivity (eg. [@r497]), and so cannot easily be utilized in large-scale multi-species planning exercises. Thirdly, we used amplified fragment length polymorphism data (AFLP; [@r453]) to characterize the main gradients of genetic variation. Whilst next-generation sequencing provides higher resolution genetic information than AFLP data ([@r480]), we know of no suitable multi-species genomic data set, and our methodology would still have utilized only the main gradients of the genetic variation to generate prioritizations in a computationally feasible period of time.

@@

We used data for 27 alpine plant species in the European Alps collected by the IntraBioDiv project ([@r451]; Figure 2a). All data, code, and results are stored in an online repository to permit replication and validation of this study (\url{www.github.com/jeffreyhanson/genetic.surrogates}). This data set has been used to explore patterns of adaptive (eg. [@r455]) and neutral genetic variation (eg. [@r478]), and the potential for species richness as a potential surrogate for genetic diversity ([@r486]). Data was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx. 22.3 km $\times$ 25 km; Figure S1). Project members visited every second grid cell, and for each species, they collected samples from three individuals in each cell they were detected in. They genotyped samples using AFLPs, and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; for more information see [@r478], Table 1). Thus the data set contains information describing the genomic properties of individuals for each species over a geographic grid. We used this data set because it comprized comparable genetic data for a range of species with different evolutionary and life histories that was collected using a standardized sampling scheme. Additionally, the study area contained large environmental variation which is useful for examining environmental surrogates.

@@

We adopted the sampling grid used to collect data as planning units to develop prioritizations. To reduce computational burden, we omitted cells that did not contain samples (`r nrow(grid.DF)` grid cells used in analysis). To examine the potential effects of variation in cost, we calculated the total human population density inside each planning unit (1 km$^2$ resolution from the Global Rural-Urban Mapping Project; [@r472]) and used this to represent acquisition cost (Figure 2b). All spatial and statistical analyses were conducted in \texttt{R} (version `r paste0(R.version$major, '.', R.version$minor)`; [@r192]).

@@

After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species. We discarded the population groupings, and partitioned species' adaptive and neutral loci into separate matrices. We applied NMDS (implemented in the \texttt{vegan R} package; [@r458]) using Gower distances (via the \texttt{cluster R} package; [@r457]) to derive continuous variables that described the main gradients of adaptive and neutral genetic variation separately for each species (Table S2). To ensure that the ordinations described a sufficient amount of the genetic variation, we ran successive scaling analyses with increasing dimensionality until a sufficient stress value was obtained (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis). Since each grid cell had multiple samples per species, we averaged the ordinated values belonging to conspecifics in the same grid cell.

@@

prioritization method {#prioritization-method .unnumbered}
----------------------------------------------------------

We used the unreliable representative and adequate prioritization (URAP) formulation for reserve selection ([@r465]). This formulation can identify the minimum number of planning units required to preserve both a target proportion of the species' geographic range (using amount-based targets) and a broad sample of variation found across the species' range. We used the environmental and geographic surrogate spaces, and the adaptive (if detected) and neutral genetic spaces to generate and evaluate solutions. Solutions associated with negative values--because they secured a minute fraction of genetic variation--were replaced with zeros to facilitate statistical analysis. We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality (using the \texttt{raptr R} package [@r465]).

@@

Our first aim was to determine if the environmental and geographic variables correlate with spatial patterns of adaptive and neutral genetic variation. To achieve this, for each species, we generated multiple dissimilarity matrices expressing differences between the planning units where the species was found. Using data from the species' surrogate and genetic spaces, these dissimilarity matrices showed differences between the planning units in terms of their (i) geographic position, (ii) environmental characteristics, and the (iii) adaptive (if detected) and (iv) neutral genetic characteristics of the individuals inside them. Dissimilarities were computed as Euclidean distances.

We then fitted maximum likelihood population-effects (MLPE) models ([@r494]). These models accommodated the structure of dissimilarity matrices using random effects. For each species, we fitted a MLPE model to the species' dissimilarity matrices based on geographic position and neutral genetic variation. Additionally, if adaptive loci were detected, we also fitted a MLPE model to the species' dissimilarity matrices describing environmental and adaptive genetic variation. All predictor variables were _z_-transformed to improve convergence. To test if the surrogates explained a substantial amount of genetic variation, we compared each model to its null model using $\chi^2$, and applied Bonferroni corrections.

>>>>

---
title: "Environmental and geographic variables are effective surrogates for genetic variation in conservation planning"
author:
  - name: Jeffrey O. Hanson
    affiliation: a,1
  - name: Jonathan R. Rhodes
    affiliation: b
  - name: Cynthia Riginos
    affiliation: a
  - name: Richard A. Fuller
    affiliation: a
address:
  - code: a
    address: School of Biological Sciences, The University of Queensland, Brisbane, QLD, Australia
  - code: b
    address: School of Geography, Planning and Environmental Management, The University of Queensland, Brisbane, QLD, Australia
corresponding_author:
  - code: 1
  - text: "To whom correspondence should be addressed. E-mail: jeffrey.hanson@uqconnect.edu.au"
author_contributions: |
  All authors developed the study and drafted the manuscript. JOH performed all analyses.
lead_author_surname: "Hanson"
abstract: |
  Protected areas buffer species from anthropogenic threats and provide places for the processes that generate and maintain biodiversity to continue. However, genetic variation, the raw material for evolution, is difficult to capture in conservation planning, not least because genetic data require considerable resources to obtain and analyze. Here we show that freely available environmental and geographic distance variables are highly effective surrogates in conservation planning for representing adaptive and neutral intra-specific genetic variation. We obtained occurrence and genetic data from the IntraBioDiv project for 27 alpine plant species collected over the European Alps using a gridded sampling scheme. For each species, we identified loci that were potentially under selection using outlier loci methods, and mapped their main gradients of adaptive and neutral genetic variation across the grid cells. We then used the cells as planning units to prioritize protected area acquisitions. First, we verified that the spatial patterns of environmental and geographic variation were correlated, respectively, with adaptive and neutral genetic variation. Second, we showed that these surrogates can predict the proportion of genetic variation secured in randomly generated solutions. Finally, we discovered that solutions based only on surrogate information secured substantial amounts of adaptive and neutral genetic variation. Our work paves the way for widespread integration of surrogates for genetic variation into conservation planning.
significance: |
  To protect biodiversity for the long term, nature reserves and other protected areas need to represent a broad range of different gentic types. However, genetic data are expensive and time consuming to obtain. Here we show that freely available environmental and geographic variables can be used as effective surrogates for genetic data in conservation planning. This means that conservation plannners can, with some confidence, design protected area systems to represent genetic diversity without investing in expensive programmes to obtain and analyze genetic data.
acknowledgements: |
  JOH is funded by an Australian Postgraduate Award (APA) scholarship. RAF has an Australian Research Council Future Fellowship. This work was supported by the Centre of Biodiversity and Conservation Science (CBCS).
keywords:
  - conservation
  - biodiversity
  - AFLP
  - genetic diversity
pnas_type: pnasresearcharticle
bibliography: references.bib
csl: reference-style.csl
output:
  rticles::pnas_article:
    includes:
      in_header: [preamble.tex]
---

<!-- NOTES -->
<!-- 250 word limit for abstract -->
<!-- 120 word limit for significance statement -->
<!-- 5000 word length limit -->

@@

```{r,include=FALSE}
# load .rda file
checkpoint::checkpoint('2016-11-26', R.version='3.3.2', scanForPackages=FALSE)
try(session::restore.session('data/final/results.rda'))
pvalString <- lazyWeave::pvalString

@@

Although the strength of natural selection is a continuous force, genetic variation is often classified as adaptive or neutral (reviewed in [@r480]). Adaptive genetic variation is associated with loci that significantly affect fitness. Typically, such "adaptive" loci are detected due to anomalous patterns among individuals (eg. [@r456; @r487]), though phenotypic-genetic associations are more likely to identify functionally important genomic regions for individual species (eg. [@r500]). By representing the full range of adaptive variation, protected areas can help ensure that populations with particularly beneficial adaptations are not lost ([@r3; @r43]). In contrast, neutral genetic variation is associated with loci that do not significantly affect fitness, but instead reflect the evolutionary history of different populations. By representing the full range of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity ([@r67]). Thus optimally sited protected area networks would aim to represent both adaptive and neutral patterns of genetic variation ([@r67; @r3; @r43]).

Recently, the use of surrogate data has been proposed to generate prioritizations that capture intra-specific genetic variation ([@r2; @r200])--without needing to utilize genetic data to achieve this. Since adaptation is ultimately driven by selection pressures, environmental variables have been proposed as surrogates for adaptive genetic variation (eg. [@r202; @r2]). Specifically, by ensuring that prioritizations secure individuals in a broad range of environmental conditions, rather than individuals in similar conditions, we would expect to capture a large amount of adaptive genetic variation. However, the effectiveness of this approach remains unverified. Neutral genetic variation arizes from a reduction in gene flow between populations. Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood (reviewed in [@r501]). Potential surrogates for neutral genetic variation have been based on variables that predict the levels of connectivity between planning units. For instance, isolation by distance (IBD; [@r259]) predicts that populations located further apart will experience less gene flow, and in turn, share fewer neutral loci. Based on this idea, spreading conservation effort evenly across the geographic distribution of an island network has been found to capture more neutral genetic variation ([@r200]). However this remains untested in spatially contiguous systems where connectivity is complicated by many additional factors (eg. habitat; [@r497])--the typical situation in most conservation planning exercises.

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=(5*0.9), fig.height=(7.1*0.9), out.width='3.42in', out.height='4in', fig.cap='The proportion of adaptive and neutral genetic variation secured for each species in solutions based on different targets and scenarios. Box plots show the 25th percentile, median, and 75th percentile. Whiskers show the data at 1.5 $\times$ the inter-quantile range. Points show data outside the whiskers. Red lines show the target proportion of genetic variation.'}
# prepare data for plotting
scenario.SDF <- scenario.DF %>%
  group_by(Prioritisation, Context, Metric, Species) %>%
  summarize(genetic.held=mean(genetic.held, na.rm=TRUE)) %>%
  data.frame()

# make plot
suppressWarnings({
  ggplot(aes(x=Prioritisation,y=genetic.held),
    data=scenario.SDF) +
    geom_boxplot(position=position_dodge(0.9)) +
    scale_fill_manual(name='Prioritization',
      values=c('grey80','grey50','grey20')) +
    geom_abline(slope=0, color='red', linetype='dashed',
      intercept=raptr.params.LST[[MODE]]$scenario.analysis$genetic.target) +
    ylab('Genetic variation secured (%)') +
    xlab('') +
    scale_y_continuous(breaks=c(0,0.25,0.5,0.75,1)) +
    theme_classic() +
    theme(strip.background = element_rect(fill='grey20'),
      strip.text = element_text(color='white', size=9),
      panel.border = element_rect(color='black', fill=NA),
      axis.text.x=element_text(angle=45, hjust=1, vjust=1, size=8),
      axis.text.y=element_text(size=8)) +
    facet_grid(Context ~ Metric)
})
```

The proportion of genetic variation secured in a prioritization depended on the targets used to generate the prioritization (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$p.value[[2]])`)$. Most notably, prioritizations based on the surrogate variables represented significantly more genetic variation than amount-based prioritizations (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` SD overall genetic variation secured versus `r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` SD; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$), and were not distinguishable from prioritizations based on measured genetic data (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` SD; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$). These results indicate that environmental and geographic surrogates in this case far outperform traditional amount-based conservation planning, and perform as well as a conservation plan based directly on measured genetic data.

@@

<!-- surrogate/genetic held correlations -->
Our results demonstrate that environmental and geographic variables are effective surrogates for most species (Figure 2). Despite this, for a few species, geographic distance was a suprisingly poor surrogate for neutral genetic variation (eg. _Gypsophila repens_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gypsophila_repens')$Geographic.r2CU,2)`; _Luzula alpinopilosa_: $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Luzula_alpinopilosa')$Geographic.r2CU,2)`; _Gentiana nivalis_ $R^2$ = `r round(filter(surrogate.spp.DF, Species=='Gentiana_nivalis')$Geographic.r2CU,2)`). One explanation for this result is that geographic distance will be a poor surrogate for neutral genetic variation where connectivity is complicated by additional factors. For example, species can maintain high connectivity between distant populations (eg. wind dispersing plants; [@r451]), and connectivity between nearby populations can be disrupted by landscape features ([@r497]). Overall, such species were the exception in our analysis, and both surrogates performed moderately well for most species.

<!-- suitable targets for surrogates -->
Our results show that environmental and geographic surrogates can be used to capture genetic variation in prioritizations, the next question is: what percentage of surrogate variation should planners target to preserve genetic variation? To secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` % of the species' neutral genetic variation, prioritizations needed to sample `r round(mean(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$geographic.held*100,na.rm=TRUE),2)` SD of the variation in geographic position among the planning units occupied by the species. Additionally, to secure at least `r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` % of each species' adaptive genetic variation, prioritizations needed to sample `r round(mean(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` % $\pm$ `r round(sd(surrogate.repr.prop.DF$environmental.held*100,na.rm=TRUE),2)` SD of the environmental variation in planning units occupied by the species. For all species, however, it was possible to generate solutions that secured a high proportion of the surrogate variation (> 90 %) and only a small proportion of genetic variation (< 20 %; Figures S56--S57). Planners may avoid such outcomes by using both amount- and surrogate-based targets. Depending on study area and the species of conservation interest, higher targets may be needed to ensure that prioritizations secure a representative sample of the genetic variation for all of the species in a planning exercise.

<!-- adaptive genetic variation secured in prioritizations-->
Our results suggest that conservation planners can secure a reasonably representative sample of the intra-specific adaptive genetic variation using only amount-based targets (Figure 2). For most species, the spatial patterns of adaptive genetic variation clustered into a few distinct groups (eg. _Carex semipervens_ and _Cerastium uniflorum_; Figures S33 and S34 respectively). As a consequence, a prioritization would only need to select a single planning unit in each cluster to secure a representative sample. This mechanism explains why adaptive variation accumulated much more quickly than neutral genetic variation in the randomly generated prioritizations. Additionally, this mechanism also explains why the single-species prioritizations generated using only amount targets were often able to achieve high representation--they were essentially a random selection and so frequently contained planning units in different parts of the species' ranges (Figure 4a). Overall--and contrary to our original hypotheses--it seems that conventional reserve selection methods might often conserve species' adaptive genetic variation.

<!--  limitations/ weaknesses -->
There are several limitations associated with our analysis. Firstly, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in regional conservation planning (1--10 km$^2$; @rXXX). We used this resolution because the genetic data were collected at this scale. Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis by artificially introducing additional spatial auto-correlation. Secondly, we used geographic distances as surrogates for neutral genetic variation, yet distances that incorperate data on topography may better describe connectivity in the study area. However, such distances often require species-specific scaling (eg. [@r497]), and so cannot easily be utilized in multi-species planning exercises. Thirdly, we used amplified fragment length polymorphism data (AFLP; [@r453]) to describe genetic variation. Whilst next-generation sequencing provides higher resolution genetic information than AFLP data ([@r480]), we know of no suitable multi-species genomic data set, and our methodology would still have utilized only the main gradients of the genetic variation to generate prioritizations in a feasible period of time.

@@

We used data for 27 alpine plant species in the European Alps collected by the IntraBioDiv project ([@r451]; Figure 2a). All data, code, and results are stored in an online repository to permit replication and validation of this study (\url{www.github.com/jeffreyhanson/genetic.surrogates}). This data set has been used to explore patterns of adaptive (eg. [@r455]) and neutral genetic variation (eg. [@r478]), and the potential for species richness as a surrogate for genetic diversity ([@r486]). Data was collected using a $20^{\prime}$ longitude by $21^{\prime}$ latitude grid (approx. 22.3 km $\times$ 25 km; Figure S1). Project members visited every second grid cell, and for each species, they collected samples from three individuals if they were detected. They genotyped samples using AFLPs, and constructed matrices denoting the presence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; for more information see [@r478], Table 1). Thus the data set contains information describing the genomic properties of individuals for each species over a geographic grid. We used this data set because it comprized comparable genetic data for a range of species with different evolutionary and life histories that was collected using a standardized sampling scheme.

@@

We adopted the sampling grid used to collect data as planning units to develop prioritizations. To reduce computational burden, we omitted cells that did not contain samples (`r nrow(grid.DF)` grid cells used in analysis). To examine the effect of variation in cost, we calculated the total human population density inside each planning unit (1 km$^2$ resolution from the Global Rural-Urban Mapping Project; [@r472]) and used this to represent acquisition cost (Figure 2b). All spatial and statistical analyses were conducted in \texttt{R} (version `r paste0(R.version$major, '.', R.version$minor)`; [@r192]).

@@

After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species. We discarded the population groupings, and partitioned species' adaptive and neutral loci into separate matrices. We applied NMDS (implemented in the \texttt{vegan R} package; [@r458]) using Gower distances (via the \texttt{cluster R} package; [@r457]) to derive continuous variables that described the main gradients of adaptive and neutral genetic variation separately for each species (Table S2). To ensure that the ordinations described a sufficient amount of the genetic variation, we ran successive scaling analyses with increasing dimensionality until a sufficient stress value was obtained (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis). Since each grid cell had multiple samples per species, we averaged the ordinated values for conspecifics in the same grid cell.

@@

Prioritization method {#prioritization-method .unnumbered}
----------------------------------------------------------

We used the unreliable representative and adequate prioritization (URAP) formulation to generate solutions ([@r465]). This formulation can identify the cheapest set of planning units required to preserve both a target proportion of the species' geographic range (using amount-based targets) and a broad sample of variation found across the species' range. We used the environmental and geographic surrogate spaces, and the adaptive (if detected) and neutral genetic spaces to generate and evaluate solutions. Solutions associated with negative values--because they secured very little genetic variation--were replaced with zeros to facilitate statistical analysis. We solved all reserve selection problems to within `r (gurobi.params.LST[[MODE]]$MIPGap*100)` % of optimality (using the \texttt{raptr R} package [@r465]).

@@

Our first aim was to determine if the environmental and geographic variables correlate with the spatial patterns of adaptive and neutral genetic variation. To achieve this, for each species, we created dissimilarity matrices using Euclidean distances and the data in the surrogate and genetic spaces. These matrices showed differences between the planning units occupied by the species in terms of the units' (i) geographic position, (ii) environmental characteristics, and the (iii) adaptive (if detected) and (iv) neutral genetic characteristics of the individuals inside them.

We fitted maximum likelihood population-effects (MLPE) models using maximum likelihood to investigate correlations between the dissimilarity models (using the \texttt{lme4 R} package; [@r469]). These models use random effects to accommodate the structure of dissimilarity matrices. For each species, we fitted a MLPE model to the species' dissimilarity matrices based on geographic position and neutral genetic variation. If adaptive loci were detected, we also fitted a MLPE model to the species' dissimilarity matrices describing environmental and adaptive genetic variation. All data variables were _z_-transformed to improve convergence. To test if the surrogates explained the genetic variation, we compared each model to its null model using a $\chi^2$ test, and applied Bonferroni corrections.

Our second aim was to determine if the environmental and geographic variables were effective surrogates for adaptive and neutral genetic variation. For each species, we generated `r format(raptr.params.LST[[MODE]]$surrogacy.analysis$random.replicates*2, big.mark=',', scientific=FALSE)` prioritizations by randomly selecting different combinations of planning units that the species occupied. For half of these random prioritizations, we calculated the proportion of geographic variation and neutral genetic variation they secured, and for the other half, we calculated the proportion of environmental variation and adaptive genetic variation they secured. The proportion of surrogate and genetic variation captured in a prioritization was calculated using the \texttt{raptr R} package.
